SAT Competition 2020\tnoteref{title}

### 2021-01-07 paper/main.tex
the remaining tables

From what was once mainly the archetypal intractable (in particular NP-complete) problem, propositional satisfiability 
(or Boolean satisfiability, SAT) has flourished in practice into a success story of modern computer science.
This is due to advances in SAT solvers, i.e., implementations of decision procedures for SAT, which form today a 
central computational tool for solving real-world problem instances of various kinds of NP-hard search and optimization problems.
With standardized inputformats, readily-available APIs for incrementalapplications,
and certified proof logging andchecking capabilities, applications of SAT solver technology
have branched from first breakthrough applications in automated planning, test pattern generation and hardware verification
to thousands of different application settings.

The success of SAT would not be without the persistent efforts of the SAT community to further develop the performance and robustness
of SAT solvers. The SAT Competition series, which a history dating back to early 90s, aims to support and provide further incentives 
for maintaining and furthering this progress. In particular,

WILL CONTINUE FROM HERE -MATTI

@@

In this competition, we initiated a new application-specific sub-track. 
The idea is to provide a large amount of benchmark instances of a specific application and to evaluate all solvers with respect to these instance in a special track. 
In this competition, we used 200 instances of the planning domain to run a planning sub-track.

@@

SAT Competition 2020 consisted of four tracks:
the Main track, Incremental Library track, Parallel track,
and the Cloud Track for massively parallel SAT solvers using up to 1024 cores. 
Furthermore, the Main track comprised of several sub-tracks.

\subsubsection{Main Track}

@@

Solvers not complying with either of the above criteria were only evaluated in a so-called No-Limits sub-track and were not eligible for the Main track awards. The No-Limits thus enabled participation of closed-source solvers (not being able or willing to expose the source code for legal or other reasons) as well as portfolio solvers (combining two or more core SAT solvers developed by different groups of authors; c.f.~Sect.~\ref{sec:rules}). However, solvers in No-Limits still competed against all other solvers in the Main Track (hence, we consider it a sub-track). The No-Limits sub-track was only evaluated with respect to the benchmarks instances which were \emph{newly} submitted to SAT Competition 2020.

Part of the Main track was the Planning sub-track, in which we evaluated the solvers on 200 benchmarks 
which all came from the same application domain -- automated planning. 
Solver submitted to the Main track automatically participated in the Planning sub-track.
It is envisioned, that also in the future there will be an application sub-track of the Main track, each time highlighting a different area where the SAT solving technology helps to advance the state of the art.

@@

\paragraph{Source Code}
The source code of submitted SAT solvers had to be made available 
(licensed for research purposes) except for the solvers participating only in the No-Limits sub-track.

@@

Apart from the No-limits sub-track, participants were not allowed to submit a portfolio of solvers,
i.e., a combination of two or more core SAT solvers developed by different groups of authors.\footnote{
In other words, a submission of a combination of solvers was only possible if all the authors of all the parts
were explicitly listed. This means that all the authors had to be notified if such participation
was planned and had to consider it carefully, also taking into account the limited number of submissions per author
as specified by the previous rule.}

@@

Solvers were ranked using a PAR-2 score based on a \num{5000}-second timeout.
A PAR-2 system assigns as many points as the number of seconds it took the solver
to solve a particular instance and twice the time limit, i.e.~\num{10000} points,
if the instance was not solved. This means that lowers scores are better.
We generally report the average score of each solver
across a benchmark set.

@@

In all tracks it was required to output a model to certify recognising a satisfiable instance.
On the other hand, certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits sub-track).

@@

\label{sec:computing}
The Main Track with its sub-tracks was run on the StarExec cluster \cite{starexec},
whose nodes are equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was \SI{5000}{\second}. 
(In the Main track, validation then continued for up to \SI{45000}{\second}.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition.}

@@

@@

\subsection{Planning Sub-Track}
% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
    % - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the planning sub-track are big compared to the rest of the
main track. Using the number of clauses as a metric; out of the $100$ biggest
instances
% of the main-track
$86$ belong to the planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances, where the average for the rest of the instances from the
main track is below $60$\%.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} and improved clause-learing based on the notion of \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful and won the first price in the Planning Sub-Track.

@@

>>>>

From what was once mainly the archetypal intractable (in particular NP-complete) problem, propositional satisfiability 
(or Boolean satisfiability, SAT) has flourished in practice into a success story of modern computer science.
This is due to advances in SAT solvers, i.e., implementations of decision procedures for SAT, which form today a 
central computational tool for solving real-world problem instances of various kinds of NP-hard search and optimization problems.
With standardized input formats, readily-available APIs for incremental applications,
and certified proof logging and checking capabilities, applications of SAT solver technology
have branched from first breakthrough applications in automated planning, test pattern generation and hardware verification
to thousands of different application settings.

The success of SAT would not be without the persistent efforts of the SAT community to further develop the performance and robustness
of SAT solvers. The SAT Competition series, which a history dating back to early 90s, aims to support and provide further incentives 
for maintaining and furthering this progress. Organized yearly as an international open event, the SAT competitions (and their variants
in the form of SAT Races and SAT Challenge) have a consistent track record of receiving tens of solver submissions yearly, submitted
by the community at large for obtaining a snapshot of the current state-of-the-art in practical SAT solving. Alongside participating solvers,
the competition invite through open calls submissions of benchmark instances representing in particular new interesting applications 
scenarios of SAT solvers. Indeed, in addition to evaluating recently developed solvers, an important aspect of the SAT competition series is
to construct on a yearly basis new benchmark sets, consisting of instances from various different application settings, which together with benchmark sets from previous years constitute a standard dataset for use in research papers and SAT solver development.

This article focuses on the 2020 instantiation of the SAT competitions. 
To this end, we provide
a detailed account of SAT Competition 2020 in terms of organizational details, competition tracks,
participating solvers, benchmarks, and the empirical results from the competition.
In terms of competition tracks, two new tracks, namely the cloud track and an application-specifc track, where introduced in 2020 
in addition to the already earlier established main, parallel, and incremental tracks; we
provide motivation and the new organizational details for both of the tracks.
In terms of solvers, we provide an overview of solving strategies and other details implemented in top-performing solvers
from the competition, complementing the individual solver descriptions available in the 2020 competition proceedings.
 As for benchmarks, we describe how the 2020 benchmark sets were constructed for each of the competition tracks, with 
an overview of the benchmarks contributed to the 2020 competition.
In terms of empirical results we provide further analysis on the results going beyond the standard rankings provided
on the SAT competition webpages (\url{https://satcompetition.github.io/2020/}). Finally, we also provide a discussion on lessons learned and ideas for future editions of SAT competitions.

The rest of this article is organized as follows. TODO -MJ

@@

In this competition, we initiated a new application-specific track. 
The idea is to provide a large amount of benchmark instances of a specific application and to evaluate all solvers with respect to these instance in a special track. 
In this competition, we used 200 instances of the planning domain to run a planning track.

@@

SAT Competition 2020 consisted of seven tracks:
Main track, No-Limits track, Planning track, ``Glucose hack'' track,
Incremental Library track, Parallel track,
and the Cloud Track for massively parallel SAT solvers. % using up to 1024 cores.

\subsubsection{The Main, No-Limits, Planning, and ``Glucose hack'' tracks}

@@

Solvers not complying with either of the above criteria were only evaluated in a so-called No-Limits track and were not eligible for the Main track awards.
The No-Limits thus enabled participation of closed-source solvers (not being able or willing to expose the source code for legal or other reasons) 
as well as portfolio solvers (combining two or more core SAT solvers developed by different groups of authors; c.f.~Sect.~\ref{sec:rules}).
However, solvers in No-Limits still competed against all other solvers submitted to the Main Track.
Thus, to deserve a mention, a No-Limits solver would need to beat all the Main Track participants.
The No-Limits track was only evaluated with respect to the benchmarks instances which were \emph{newly} submitted to SAT Competition 2020.

The same rules for participation as in Main track applied to the Planning track, in which we evaluated the solvers on 200 benchmarks 
which all came from the same application domain -- automated planning. 
Solver submitted to the Main track automatically participated in the Planning track.
It is envisioned that also in the future there will be an ``application promotion'' track,
each time highlighting a different area where the SAT solving technology helps to advance the state of the art.

@@

\paragraph{Source Code}
The source code of submitted SAT solvers had to be made available 
(licensed for research purposes) except for the solvers participating only in the No-Limits track.

@@

Apart from the No-limits track, participants were not allowed to submit a portfolio of solvers,
i.e., a combination of two or more core SAT solvers developed by different groups of authors.\footnote{
In other words, a submission of a combination of solvers was only possible if all the authors of all the parts
were explicitly listed. This means that all the authors had to be notified if such participation
was planned and had to consider it carefully, also taking into account the limited number of submissions per author
as specified by the previous rule.}

@@

Solvers were ranked using a PAR-2 score based on a \num{5000}-second timeout.
A PAR-2 system assigns as many points as the amount of time (in seconds) it took the solver
to solve a particular instance and twice the time limit, i.e.~\num{10000} points,
if the instance was not solved. This means that lowers scores are better.
We generally report the \emph{average score} of each solver across a benchmark set.

@@

In all tracks it was required to output a model to certify recognising a satisfiable instance.
On the other hand, certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits track).

@@

\label{sec:computing}
The Main, No-Limits, and Planning tracks were run on the StarExec cluster \cite{starexec},
whose nodes are equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was \SI{5000}{\second}. 
(In the Main track, validation then continued for up to \SI{45000}{\second}.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition.}

@@

\begin{table}
\caption{Main ALL}
\label{tab:mainALL}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
1  & 3926.2 & 264 & {\bf Kissat-sc2020-sat} \\
(1)& 4083.1 & 260 & Kissat-sc2020-default \\
2  & 4179.3 & 253 & {\bf Relaxed\_LCMDCBDL\_newTech} \\
3  & 4266.7 & 248 & {\bf cryptominisat-ccnr-lsids} \\
(3)& 4278.0 & 250 & cryptominisat-ccnr \\
   & 4428.1 & 250 & cadical-alluip-trail \\
   & 4429.6 & 250 & cadical-alluip \\
   & 4436.5 & 245 & Relaxed\_LCMDCBDL \\
   & 4501.2 & 243 & cryptominisat-walksat \\
   & 4554.0 & 243 & cadical-trail \\
   & 4560.0 & 238 & Kissat-sc2020-unsat \\
   & 4607.6 & 246 & CaDiCaL-sc2020 \\
   & 5157.7 & 216 & Undominated-LC-MapleLCMDiscChronoBT-DL \\
   & 5159.6 & 220 & MapleLCMDistChronoBT-DL-f2trc \\
   & 5177.0 & 214 & ExMapleLCMDistChronoBT\_\_PADC\_DL \\   
\end{tabular}
\end{table}

\begin{table}
\caption{Main SAT}
\label{tab:mainSAT}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 2997.4 & 150 & {\bf Relaxed\_LCMDCBDL\_newTech} \\
 2 & 3127.6 & 146 & {\bf Kissat-sc2020-sat} \\
 3 & 3263.0 & 144 & {\bf cryptominisat-ccnr-lsids} \\
(3)& 3317.4 & 145 & cryptominisat-ccnr \\
   & 3355.5 & 143 & Relaxed\_LCMDCBDL \\
   & 3524.7 & 140 & cryptominisat-ccnr-nolimits \\
   & 3622.4 & 138 & cryptominisat-ccnr-lsids-nolimits \\
   & 3703.7 & 140 & cryptominisat-walksat-nolimits \\
   & 3721.2 & 139 & cryptominisat-walksat \\
   & 3830.5 & 134 & Kissat-sc2020-default \\
   & 3908.5 & 135 & cadical-alluip-trail \\
   & 3909.6 & 135 & cadical-alluip \\
   & 4265.6 & 126 & cadical-trail \\
   & 4368.6 & 130 & CaDiCaL-sc2020 \\
   & 4804.9 & 114 & Kissat-sc2020-unsat \\
\end{tabular}
\end{table}

\begin{table}
\caption{Main UNSAT}
\label{tab:mainUNSAT}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
1  & 4315.1 & 124 & {\bf Kissat-sc2020-unsat} \\
(1)& 4335.6 & 126 & Kissat-sc2020-default \\
(1)& 4724.8 & 118 & Kissat-sc2020-sat \\
 2 & 4842.5 & 117 & {\bf cadical-trail} \\
-- & 4846.7 & 116 & CaDiCaL-sc2020 \\
(2)& 4947.8 & 115 & cadical-alluip-trail \\
(2)& 4949.6 & 115 & cadical-alluip \\
 3 & 4991.4 & 110 & {\bf MapleLCMDistChronoBT-f2trc-s} \\
(3)& 5051.4 & 109 & MapleLCMDistChronoBT-f2trc \\
(3)& 5054.3 & 110 & MapleLCMDistChronoBT-DL-f2trc \\
   & 5069.6 & 109 & Maple\_mix \\
   & 5110.0 & 107 & ExMapleLCMDistChronoBT\_\_PADC\_DL \\
   & 5126.9 & 107 & Maple\_simp \\
   & 5128.9 & 107 & MapleLCMDistChronoBT-DL-v3 \\
   & 5142.2 & 106 & Undominated-LC-MapleLCMDiscChronoBT-DL \\
\end{tabular}
\end{table}

\begin{table}
\caption{Planning (ALL)}
\label{tab:planning}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 6406.9 & 80 & {\bf cadical-alluip-trail} \\
(1)& 6409.3 & 80 & cadical-alluip \\
 2 & 6466.9 & 79 & {\bf cryptominisat-ccnr-lsids} \\
(2)& 6471.9 & 79 & cryptominisat-ccnr \\
(2)& 6472.9 & 79 & cryptominisat-walksat \\
 3 & 6596.4 & 75 & {\bf Kissat-sc2020-unsat} \\
   & 6650.1 & 79 & cadical-trail \\
   & 6713.0 & 75 & Maple\_mix \\
   & 6746.5 & 75 & MapleCOMSPS\_LRB\_VSIDS\_2\_\_LRB\_VSIDS\_2\_init \\
   & 6754.3 & 73 & Maple\_simp \\
   & 6773.0 & 73 & MapleCOMSPS\_LRB\_VSIDS\_2\_\_default\_drup \\
   & 6773.9 & 73 & Kissat-sc2020-sat \\
   & 6774.2 & 73 & Kissat-sc2020-default \\
   & 7003.0 & 69 & optsat\_m20 \\
   & 7025.5 & 68 & upGlucose-3.0\_PADC \\
\end{tabular}
\end{table}

\begin{table}
\caption{Parallel (ALL)}
\label{tab:parallelALL}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 3316.6 & 283 & {\bf pmcompspsthirty} \\
(1)& 3714.7 & 271 & pmcompspsixty \\
 2 & 3743.4 & 269 & {\bf plingeling} \\
 3 & 3985.3 & 260 & {\bf manyglucosethirty} \\
   & 4022.7 & 262 & painlessmaplevone \\
   & 4036.3 & 258 & manyglucosesixty \\
   & 4103.3 & 260 & painlessmaplevtwo \\
   & 4433.3 & 243 & syrupscavel \\
   & 4903.4 & 225 & treengeling \\
   & 5240.1 & 213 & abcdsatptwenty \\
\end{tabular}
\end{table}

\begin{table}
\caption{Parallel (SAT)}
\label{tab:parallelSAT}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 2853.7 & 153 & {\bf pmcompspsthirty} \\
 2 & 2913.6 & 154 & {\bf painlessmaplevone} \\
(2)& 3082.7 & 151 & painlessmaplevtwo \\
(1)& 3196.9 & 148 & pmcompspsixty \\
 3 & 3805.9 & 133 & {\bf plingeling} \\
   & 4048.3 & 130 & manyglucosesixty \\
   & 4076.2 & 130 & manyglucosethirty \\
   & 4675.4 & 119 & syrupscavel \\
   & 4907.1 & 114 & treengeling \\
   & 6337.5 & 85 & abcdsatptwenty \\
\end{tabular}
\end{table}

\begin{table}
\caption{Parallel (UNS)}
\label{tab:parallelUNS}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 3680.8 & 136 & {\bf plingeling} \\
 2 & 3779.5 & 130 & {\bf pmcompspsthirty} \\
 3 & 3894.3 & 130 & {\bf manyglucosethirty} \\
(3)& 4024.2 & 128 & manyglucosesixty \\
   & 4142.8 & 128 & abcdsatptwenty \\
   & 4191.1 & 124 & syrupscavel \\
   & 4232.5 & 123 & pmcompspsixty \\
   & 4899.8 & 111 & treengeling \\
   & 5123.9 & 109 & painlessmaplevtwo \\
   & 5131.9 & 108 & painlessmaplevone \\
\end{tabular}
\end{table}

\begin{table}
\caption{Claud (ALL)}
\label{tab:claud}
\begin{tabular}{cccl}
place & PAR-2 score & num. solved & solver \\
\hline
 1 & 2603.8 & 299 & {\bf mallobecs} \\
 2 & 3146.8 & 278 & {\bf toposat} \\
 3 & 4797.6 & 213 & {\bf slime} \\
   & 6800.8 & 132 & paracooba \\
   & 7299.0 & 110 & ctsat \\
   & 8468.3 & 62 & paracoobamarch \\
\end{tabular}
\end{table}

@@

\subsection{Planning Track}
% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
    % - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the planning track are big compared to the rest of the
main track. Using the number of clauses as a metric; out of the $100$ biggest
instances
% of the main-track
$86$ belong to the planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances, where the average for the rest of the instances from the
main track is below $60$\%.

@@

Based on \solver{CaDiCaL}, its variants \solver{Trail} and \solver{AllUip} present implementations of a new \emph{Trail Saving} approach~\cite{Hickey:2020:TrailSaving} and improved clause-learing based on the notion of \emph{Stable AllUIP}~\cite{Bacchus:SC2020}. 
Submitted were the three variants \solver{Trail}, \solver{AllUip} and \solver{AllUip+Trail}. 
The variants including \emph{Stable AllUIP} were the most successful and won the first price in the Planning Track.

@@

\section{More detailed analysis of the Main track results (to be renamed)}

In this section, we provide additional views on the Main track results,
in particular focusing on metrics which help to highlight strengths

\subsection{Contributions to VBS}

Everything could be done separately for (ALL, SAT, UNSAT)

3X VBS

Greedy cover

Fair split strategy

Z3 strategy

Maxent Nash

Van Gelder ?
